---
title: "E1 — Corpus Descriptives (Study 1)"
author: "EduStories / habilitace-2"
date: today
format:
  html:
    embed-resources: true
    toc: true
execute:
  warning: false
  message: false
---

Study 1 (PLAN.md §5 E1): descriptive profile of the full corpus snapshot S6
(May 2026), corpus growth across named snapshots, publication drift (S6b),
and the human-coded K1 descriptives (behaviours / solutions / outcomes) on the
reproducible `K1_rez` slice — with the pre-specified DUAL counts (cases with
category vs total assignments). Text-length statistics conservatively exclude
all 52 cases ever affected by extraction truncation (8 of them were later
repaired but stay excluded; D4). Writes `vystupy/tabulky/kap5_korpus_cisla.csv`.

```{r setup}
suppressPackageStartupMessages({
  library(dplyr); library(tidyr); library(readr); library(stringr)
})
SEED <- 20260714; set.seed(SEED)
H2 <- normalizePath(file.path(getwd(), "../.."))
S6  <- file.path(H2, "data/raw/korpus/export-2026-05/Edustories_export_may26_svabik_v2.csv")
S6B <- file.path(H2, "data/raw/korpus/export-2026-05/Edustories_export_may26_svabik.xlsx")
REZ <- file.path(H2, "data/processed/k1_rez_ids.csv")
TRUNC <- file.path(H2, "data/processed/truncated_uids.csv")
sha256 <- function(p) strsplit(system2("shasum", c("-a", "256", shQuote(p)),
                                       stdout = TRUE), " ")[[1]][1]
S6_SHA <- sha256(S6); REZ_SHA <- sha256(REZ)
manifest <- list()
put <- function(metric, value, sha = S6_SHA) manifest[[length(manifest) + 1]] <<-
  data.frame(metric = metric, value = round(as.numeric(value), 4), data_sha = sha)
```

## Full corpus S6 (May 2026)

```{r s6}
s6 <- read_csv(S6, col_types = cols(.default = col_character()))
put("n_korpus_s6", nrow(s6))
put("n_zverejneno_s6", sum(s6$is_published == "true", na.rm = TRUE))
s6$uid <- paste0("exp", s6$id)
trunc <- read_csv(TRUNC, show_col_types = FALSE)
s6_len <- s6 |> filter(!(uid %in% trunc$case_uid))
put("n_vyrazeno_delky_trunc", nrow(s6) - nrow(s6_len))
wc <- function(x) str_count(x, "\\S+")
len_rows <- lapply(c("description", "anamnesis", "solution", "outcome"), function(part) {
  w <- wc(s6_len[[paste0(part, "_cs")]])
  put(paste0(part, "_slov_median"), median(w, na.rm = TRUE))
  put(paste0(part, "_slov_q1"), quantile(w, .25, na.rm = TRUE))
  put(paste0(part, "_slov_q3"), quantile(w, .75, na.rm = TRUE))
  data.frame(cast = part, median = median(w, na.rm = TRUE),
             q1 = quantile(w, .25, na.rm = TRUE), q3 = quantile(w, .75, na.rm = TRUE))
})
knitr::kable(bind_rows(len_rows), digits = 0, row.names = FALSE,
             caption = "Word counts (Czech text; all 52 truncation-affected cases excluded)")
age <- suppressWarnings(as.numeric(s6$age))
# hygiena: věk 0 = nevyplněno (kód formuláře), ne skutečný věk → NA;
# horní mez neořezáváme (žák SŠ může být 20+), počet implauzibilních dokládáme
put("vek_n_nula_vyrazeno", sum(age == 0, na.rm = TRUE))
age[age == 0] <- NA
put("vek_median", median(age, na.rm = TRUE)); put("vek_n", sum(!is.na(age)))
ai_cols <- grep("^ai_solution_.*_cs$", names(s6), value = TRUE)
put("n_ai_reseni_kompletni", sum(rowSums(!is.na(s6[ai_cols]) & s6[ai_cols] != "") == 6))
```

## Corpus growth and publication drift

Named snapshots (documented in DATA_README §1); S6b is the older export cut —
its cases missing from S6 were unpublished in the meantime.

```{r growth}
snap <- data.frame(
  rez = c("S1 (2023)", "S3 (04/2024)", "S4 (08/2024)", "S5 HF (release)", "S6 (05/2026)"),
  n = c(757, 1507, 1695, 1492, nrow(s6)))
knitr::kable(snap, row.names = FALSE)
for (i in seq_len(nrow(snap))) put(paste0("rust_", gsub("[^A-Za-z0-9]", "", snap$rez[i])), snap$n[i])
s6b_ids <- suppressMessages(readxl::read_excel(S6B, sheet = "out"))$id
s6b_ids <- as.integer(na.omit(s6b_ids))
drift <- sum(!(s6b_ids %in% as.integer(s6$id)))
put("n_s6b", length(s6b_ids), sha256(S6B))
put("n_odpublikovano_s6b_vs_s6", drift, sha256(S6B))
cat("S6b:", length(s6b_ids), "případů; z nich", drift, "už není v S6 (odpublikováno).\n")
```

## K1 human coding on the `K1_rez` slice — DUAL counts

Multi-label: a case may carry several categories. We therefore report BOTH
„cases with category" and „assignments" (PLAN §5 ⊕); their ratio explains why
assignment totals exceed the number of cases.

```{r k1rez}
rez <- read_csv(REZ, show_col_types = FALSE)
put("n_k1_rez", nrow(rez), REZ_SHA)
# normalizace štítků: sjednotit velikost písmen a známé překlepy na kanonickou formu
canon_label <- function(p) {
  p <- gsub("Částěčný", "Částečný", p)
  p <- gsub("odbonríky", "odborníky", p)           # překlep v datech
  substr(p, 1, 1) <- toupper(substr(p, 1, 1))       # „rozhovor" → „Rozhovor"
  p
}
split_cats <- function(x) {
  x <- x[!is.na(x)]
  unlist(lapply(x, function(v) {
    p <- trimws(unlist(strsplit(v, "[;,]")))
    p <- canon_label(p)
    p[p != "" & !(p %in% c("-", "?", "(zvolit)"))]
  }))
}
dual <- lapply(c(chovani = "problems_annotated", reseni = "solutions_annotated",
                 dopad = "implications_annotated"), function(col) NULL)
for (nm in names(c(chovani = 1, reseni = 1, dopad = 1))) {
  col <- c(chovani = "problems_annotated", reseni = "solutions_annotated",
           dopad = "implications_annotated")[[nm]]
  cats_per_case <- lapply(rez[[col]], function(v) unique(split_cats(v)))
  assignments <- unlist(cats_per_case)
  n_cases_with <- table(assignments)                 # případů s kategorií
  n_assign <- length(assignments)
  put(paste0("k1_", nm, "_prirazeni_celkem"), n_assign, REZ_SHA)
  put(paste0("k1_", nm, "_pripadu_s_udajem"), sum(lengths(cats_per_case) > 0), REZ_SHA)
  put(paste0("k1_", nm, "_pomer_prirazeni_na_pripad"),
      n_assign / sum(lengths(cats_per_case) > 0), REZ_SHA)
  top <- sort(n_cases_with, decreasing = TRUE)[1:min(8, length(n_cases_with))]
  for (k in names(top))
    put(paste0("k1_", nm, "_pripadu:", substr(k, 1, 38)), top[[k]], REZ_SHA)
  # vzácné strategie: explicitní metrika pro „proaktivní" (tvrzení o řídkosti v kap. 5)
  if (nm == "reseni" && "Proaktivní řešení" %in% names(n_cases_with))
    put("k1_reseni_pripadu:Proaktivní řešení", n_cases_with[["Proaktivní řešení"]], REZ_SHA)
  print(knitr::kable(data.frame(kategorie = names(top), pripadu = as.integer(top)),
                     caption = paste0(nm, " — top kategorie (počet PŘÍPADŮ)")))
}
```

## Distributions for book figures (word lengths, age)

Case-level aggregates for `fig_05_*.R` (no case texts — safe to publish).

```{r rozdeleni}
TABDIR <- file.path(H2, "vystupy/tabulky")
delky <- bind_rows(lapply(c("description", "anamnesis", "solution", "outcome"),
  function(part) data.frame(cast = part, slova = wc(s6_len[[paste0(part, "_cs")]]))))
delky <- delky[!is.na(delky$slova) & delky$slova > 0, ]
write.csv(delky, file.path(TABDIR, "kap5_delky_rozdeleni.csv"), row.names = FALSE)
put("delka_p90_description",
    quantile(delky$slova[delky$cast == "description"], .9, na.rm = TRUE))
vek_tab <- as.data.frame(table(vek = round(age)), responseName = "n")
write.csv(vek_tab, file.path(TABDIR, "kap5_vek_rozdeleni.csv"), row.names = FALSE)
put("vek_q1", quantile(age, .25, na.rm = TRUE))
put("vek_q3", quantile(age, .75, na.rm = TRUE))
cat("rozdělení: délky", nrow(delky), "řádků; věk", nrow(vek_tab), "hodnot\n")
```

## K1 full category tables, co-occurrence and outcome-by-solution

All categories per dimension (figures), conditional co-occurrence
P(solution | behaviour) on K1_rez cases, and share of long-term success by
solution type. Unit = CASE (multi-label: a case counts once per category it
carries). Descriptive only — no causal claims.

```{r k1-hloubka}
cats_of <- function(col) lapply(rez[[col]], function(v) unique(split_cats(v)))
CH <- cats_of("problems_annotated")
RE <- cats_of("solutions_annotated")
DO <- cats_of("implications_annotated")

# (a) úplné tabulky kategorií pro figury
plne <- bind_rows(
  data.frame(dimenze = "chovani", kategorie = unlist(CH)),
  data.frame(dimenze = "reseni",  kategorie = unlist(RE)),
  data.frame(dimenze = "dopad",   kategorie = unlist(DO))) |>
  count(dimenze, kategorie, name = "pripadu") |>
  arrange(dimenze, desc(pripadu))
write.csv(plne, file.path(TABDIR, "kap5_k1_kategorie.csv"), row.names = FALSE)

# (b) ko-výskyt: P(řešení | chování), jen buňky s n_chování >= 50
beh_n <- table(unlist(lapply(CH, unique)))
kov <- do.call(rbind, lapply(names(beh_n)[beh_n >= 50], function(b) {
  idx <- vapply(CH, function(s) b %in% s, logical(1))
  res_tab <- table(unlist(RE[idx]))
  data.frame(chovani = b, reseni = names(res_tab),
             n_spolu = as.integer(res_tab),
             n_chovani = sum(idx),
             podil = as.numeric(res_tab) / sum(idx))
}))
write.csv(kov, file.path(TABDIR, "kap5_kovyskyt.csv"), row.names = FALSE)
for (i in seq_len(nrow(kov)))
  put(paste0("kov_podil:", substr(kov$chovani[i], 1, 24), "|",
             substr(kov$reseni[i], 1, 24)), kov$podil[i], REZ_SHA)

# (c) dopad podle typu řešení: podíl dlouhodobého úspěchu (jen n >= 30)
dop <- do.call(rbind, lapply(names(table(unlist(RE))), function(r) {
  idx <- vapply(RE, function(s) r %in% s, logical(1)) &
         vapply(DO, length, integer(1)) > 0
  if (sum(idx) < 30) return(NULL)
  du  <- mean(vapply(DO[idx], function(s) "Dlouhodobý úspěch" %in% s, logical(1)))
  neu <- mean(vapply(DO[idx], function(s) "Neúspěch" %in% s, logical(1)))
  data.frame(reseni = r, n = sum(idx), podil_DU = du, podil_NEU = neu)
}))
dop <- dop[order(-dop$podil_DU), ]
write.csv(dop, file.path(TABDIR, "kap5_dopad_podle_reseni.csv"), row.names = FALSE)
for (i in seq_len(nrow(dop))) {
  put(paste0("dopadDU:", substr(dop$reseni[i], 1, 30)), dop$podil_DU[i], REZ_SHA)
  put(paste0("dopadN:",  substr(dop$reseni[i], 1, 30)), dop$n[i], REZ_SHA)
}
knitr::kable(dop, digits = 3,
             caption = "Podíl dlouhodobého úspěchu podle typu řešení (deskriptivně)")
```

## Freeze: write manifest

```{r manifest}
out <- bind_rows(manifest) |>
  mutate(skript = "analyzy/notebooks/10_korpus.qmd", seed = SEED,
         datum = format(Sys.Date())) |>
  select(metric, value, skript, data_sha, seed, datum)
write.csv(out, file.path(H2, "vystupy/tabulky/kap5_korpus_cisla.csv"), row.names = FALSE)
cat("Manifest:", nrow(out), "metrik → vystupy/tabulky/kap5_korpus_cisla.csv\n")
```
