---
title: "E3 — AI vs Teacher Solutions (Study 3, K2 campaign)"
author: "EduStories / habilitace-2"
date: today
format:
  html:
    embed-resources: true
    toc: true
execute:
  warning: false
  message: false
---

Formalises the completed E3 analysis (PLAN.md §5, frozen v1.0). Authoritative
data = OSF package of the AIED manuscript (`data/externi/osf-aied-ndtkz/`),
column `Reseni_gen_AI`: 0 = AI, 1 = teacher. This notebook (a) reproduces the
manuscript's LMMs (Table 3), (b) runs the pre-specified CLMM sensitivity,
(c) computes inter-rater agreement with cluster-bootstrap 95% CIs, and
(d) writes the numbers manifest `vystupy/tabulky/kap7_k2_shoda_cisla.csv`.

```{r setup}
suppressPackageStartupMessages({
  library(dplyr); library(tidyr); library(lmerTest); library(ordinal); library(irr); library(digest)
})
SEED <- 20260714
set.seed(SEED)
ROOT <- normalizePath(file.path(getwd(), ".."))          # analyzy/
H2   <- normalizePath(file.path(ROOT, ".."))             # habilitace-2/
CSV  <- file.path(H2, "data/externi/osf-aied-ndtkz/anotace_final_data_bezdiakritiky.csv")
PUB  <- file.path(H2, "data/processed/kap7_hodnoceni_pseudo.csv")
DATA_SHA <- "ba464aaa3aa7223ba6a14009ec38dfb576db9af250079ed821bb6f9992478927"
# Studie 3 potřebuje k výpočtu jen HODNOCENÍ, ne texty kazuistik. Když je surový
# OSF soubor přítomen (autor), pseudonymizujeme anotátorky, ponecháme jen
# analytické sloupce (BEZ textů Popis/Anamneza/Řešení) a zapíšeme veřejnou
# tabulku; když chybí (Binder), čteme ji. Náhrada jmen 1:1 nemění žádný výsledek
# (anotátorka vstupuje jen jako náhodný efekt), texty korpusu zůstávají privátní.
KEEP <- c("Id", "anotator_name", "Otazka", "Hodnoceni", "Reseni_gen_AI",
          "Pristup.I", "Pristup.II", "Pristup.III")
df <- read.csv(PUB, header = TRUE, sep = ",", encoding = "UTF-8")
DATA_SHA <- digest::digest(file = PUB, algo = "sha256")
stopifnot(!anyDuplicated(df[c("Id", "anotator_name", "Otazka", "Reseni_gen_AI")]))
stopifnot(length(unique(df$Id)) == 324,
          all(sort(unique(na.omit(df$Hodnoceni))) == -2:2))
dims <- c(vhodnost     = "Jak.moc.je.vhodna.toto.reseni.pro.danou.situaci.",
          reaktivni    = "Jak.moc.je.popisovane.reseni.reaktivni.nebo.proaktivni.",
          humanisticke = "Jak.moc.je.popisovane.reseni.humanisticky.nebo.behavioralne.orientovane.",
          systemove    = "Jak.moc.je.popisovane.reseni.systemove.nebo.situacni.")
manifest <- list()
put <- function(metric, value) manifest[[length(manifest) + 1]] <<-
  data.frame(metric = metric, value = round(as.numeric(value), 4))
put("n_kazuistik", length(unique(df$Id)))
put("n_anotatorek", length(unique(df$anotator_name)))
put("n_odpovedi_dimenze", nrow(df))
put("n_missing_dimenze", sum(is.na(df$Hodnoceni)))
```

## Descriptives by source

```{r descriptives}
desc <- df |>
  filter(!is.na(Hodnoceni)) |>
  mutate(dim = names(dims)[match(Otazka, dims)],
         src = ifelse(Reseni_gen_AI == 0, "AI", "teacher")) |>
  group_by(dim, src) |>
  summarise(n = n(), mean = mean(Hodnoceni), sd = sd(Hodnoceni), .groups = "drop")
knitr::kable(desc, digits = 2)
for (i in seq_len(nrow(desc)))
  put(paste0("mean_", desc$src[i], "_", desc$dim[i]), desc$mean[i])
```

## Primary models: LMM (manuscript reproduction)

```{r lmm}
lmm_rows <- lapply(names(dims), function(nm) {
  d <- df[df$Otazka == dims[[nm]], ]
  m <- lmer(Hodnoceni ~ Reseni_gen_AI + (1 | Id) + (1 | anotator_name),
            data = d, REML = FALSE)
  fe <- summary(m)$coefficients
  ci <- confint(m, parm = "beta_", method = "Wald")
  put(paste0("lmm_b0_AI_", nm), fe[1, 1])
  put(paste0("lmm_b1_ucitel_", nm), fe[2, 1])
  put(paste0("lmm_b1_se_", nm), fe[2, 2])
  put(paste0("lmm_b1_ci_lo_", nm), ci[2, 1])
  put(paste0("lmm_b1_ci_hi_", nm), ci[2, 2])
  data.frame(dim = nm, b0_AI = fe[1, 1], b1_teacher = fe[2, 1],
             se = fe[2, 2], ci_lo = ci[2, 1], ci_hi = ci[2, 2], p = fe[2, 5])
})
knitr::kable(bind_rows(lmm_rows), digits = 3)
```

Expected (AIED Table 3): appropriateness −0.85; reactive −1.01; humanistic +0.77;
systemic −0.17 — reproduced exactly.

## Sensitivity: CLMM (pre-specified)

```{r clmm}
clmm_rows <- lapply(names(dims), function(nm) {
  d <- df[df$Otazka == dims[[nm]] & !is.na(df$Hodnoceni), ]
  m <- clmm(factor(Hodnoceni, ordered = TRUE) ~ Reseni_gen_AI +
              (1 | Id) + (1 | anotator_name), data = d)
  co <- summary(m)$coefficients
  b <- co[grep("Reseni_gen_AI", rownames(co)), , drop = FALSE]
  put(paste0("clmm_b1_ucitel_", nm), b[1, 1])
  put(paste0("clmm_or_ucitel_", nm), exp(b[1, 1]))
  # Wald 95% CI na škále OR (pro tabulku 7.2 a figuru)
  lo <- b[1, 1] - 1.96 * b[1, 2]; hi <- b[1, 1] + 1.96 * b[1, 2]
  put(paste0("clmm_or_ci_lo_", nm), exp(lo))
  put(paste0("clmm_or_ci_hi_", nm), exp(hi))
  data.frame(dim = nm, b1_teacher = b[1, 1], SE = b[1, 2],
             OR = exp(b[1, 1]), OR_lo = exp(lo), OR_hi = exp(hi), p = b[1, 4])
})
knitr::kable(bind_rows(clmm_rows), digits = 3)
```

Sensitivity check: CLMM must agree in direction with LMM for all four dimensions.

## Sensitivity: LMM with a random slope for source by case

The design is paired within case (both solutions of the same case are rated),
so the source effect may vary across cases. The random-intercept-only model
assumes a constant contrast; here we relax it with (1 + source | Id) and check
whether the fixed effect and its CI move (opponent finding, 2026-07-17).

```{r lmm-slope}
slope_rows <- lapply(names(dims), function(nm) {
  d <- df[df$Otazka == dims[[nm]], ]
  m <- suppressWarnings(
    lmer(Hodnoceni ~ Reseni_gen_AI + (1 + Reseni_gen_AI | Id) +
           (1 | anotator_name), data = d, REML = FALSE))
  fe <- summary(m)$coefficients
  ci <- suppressWarnings(confint(m, parm = "beta_", method = "Wald"))
  put(paste0("lmm_slope_b1_ucitel_", nm), fe[2, 1])
  put(paste0("lmm_slope_ci_lo_", nm), ci[2, 1])
  put(paste0("lmm_slope_ci_hi_", nm), ci[2, 2])
  put(paste0("lmm_slope_singular_", nm), as.integer(isSingular(m)))
  data.frame(dim = nm, b1_teacher = fe[2, 1],
             ci_lo = ci[2, 1], ci_hi = ci[2, 2], singular = isSingular(m))
})
knitr::kable(bind_rows(slope_rows), digits = 3)
```

## Missingness by source, including structurally absent rows

Blinding hides explicit source labels, not all textual source cues. It does not
establish ignorable missingness. Separate missing ratings within observed rows
from absent rows in the intended case × rater × dimension × source grid.

```{r missing-by-source}
mis <- df |>
  mutate(src = ifelse(Reseni_gen_AI == 0, "AI", "ucitel")) |>
  group_by(src) |>
  summarise(n = n(), n_na = sum(is.na(Hodnoceni)),
            podil_na = mean(is.na(Hodnoceni)), .groups = "drop")
knitr::kable(mis, digits = 4)
for (i in seq_len(nrow(mis)))
  put(paste0("missing_podil_", mis$src[i]), mis$podil_na[i])
rating_grid <- tidyr::expand_grid(Id = unique(df$Id), anotator_name = unique(df$anotator_name),
                                Otazka = unname(dims), Reseni_gen_AI = c(0, 1)) |>
  left_join(mutate(df, observed_row = TRUE), by = c("Id", "anotator_name", "Otazka", "Reseni_gen_AI"))
put("n_ocekavanych_bunek", nrow(rating_grid))
put("n_chybejicich_radku", sum(is.na(rating_grid$observed_row)))
put("n_chybejicich_bunek_celkem", sum(is.na(rating_grid$Hodnoceni)))
missing_grid <- rating_grid |>
  group_by(Reseni_gen_AI, Otazka, anotator_name) |>
  summarise(expected = n(), absent = sum(is.na(observed_row)),
            observed_missing = sum(!is.na(observed_row) & is.na(Hodnoceni)), .groups = "drop")
write.csv(missing_grid, file.path(H2, "vystupy/tabulky/kap7_chybeni.csv"), row.names = FALSE)
knitr::kable(missing_grid)
```

## Post-audit sensitivity: complete AI–teacher pairs per rater and dimension

```{r complete-pair-sensitivity}
complete_keys <- df |>
  group_by(Id, anotator_name, Otazka) |>
  summarise(complete = n_distinct(Reseni_gen_AI) == 2 && all(!is.na(Hodnoceni)), .groups = "drop") |>
  filter(complete)
paired_ratings <- semi_join(df, complete_keys, by = c("Id", "anotator_name", "Otazka"))
paired_results <- bind_rows(lapply(names(dims), function(dimension) {
  selected <- paired_ratings[paired_ratings$Otazka == dims[[dimension]], ]
  fitted <- lmer(Hodnoceni ~ Reseni_gen_AI + (1 | Id) + (1 | anotator_name), data = selected, REML = FALSE)
  estimate <- summary(fitted)$coefficients["Reseni_gen_AI", "Estimate"]
  interval <- confint(fitted, parm = "Reseni_gen_AI", method = "Wald")
  put(paste0("paired_n_", dimension), nrow(selected))
  put(paste0("paired_b1_", dimension), estimate)
  put(paste0("paired_ci_lo_", dimension), interval[1, 1])
  put(paste0("paired_ci_hi_", dimension), interval[1, 2])
  data.frame(dimension = dimension, n = nrow(selected), estimate = estimate,
             ci_lo = interval[1, 1], ci_hi = interval[1, 2])
}))
write.csv(paired_results, file.path(H2, "vystupy/tabulky/kap7_uplne_dvojice.csv"), row.names = FALSE)
knitr::kable(paired_results, digits = 4)
```

## Inter-rater agreement with cluster-bootstrap 95% CIs

Unit = (Id × source); Krippendorff's alpha (ordinal); bootstrap resamples the
324 case Ids with replacement (B = 1000, seed `r SEED`).

```{r alpha}
make_wide <- function(d) {
  w <- d |> mutate(item = paste0(Id, "_", Reseni_gen_AI)) |>
    select(anotator_name, item, Hodnoceni) |>
    distinct(anotator_name, item, .keep_all = TRUE) |>
    pivot_wider(names_from = item, values_from = Hodnoceni)
  m <- as.matrix(w[, -1]); rownames(m) <- w[[1]]
  m[, colSums(!is.na(m)) >= 2, drop = FALSE]
}
alpha_of <- function(m) suppressWarnings(irr::kripp.alpha(m, method = "ordinal")$value)

B <- 1000
ids <- unique(df$Id)
alpha_rows <- lapply(names(dims), function(nm) {
  d <- df[df$Otazka == dims[[nm]] & !is.na(df$Hodnoceni), ]
  a <- alpha_of(make_wide(d))
  full_matrix <- make_wide(d)
  columns_by_id <- lapply(as.character(ids), function(case_id) {
    which(colnames(full_matrix) %in% paste0(case_id, "_", c(0, 1)))
  })
  boots <- vapply(seq_len(B), function(b) {
    selected_cases <- sample(seq_along(ids), replace = TRUE)
    selected_columns <- unlist(columns_by_id[selected_cases], use.names = FALSE)
    alpha_of(full_matrix[, selected_columns, drop = FALSE])
  }, numeric(1))
  ci <- quantile(boots, c(.025, .975), na.rm = TRUE)
  put(paste0("alpha_ord_final_", nm), a)
  put(paste0("alpha_ci_lo_", nm), ci[1])
  put(paste0("alpha_ci_hi_", nm), ci[2])
  data.frame(dim = nm, alpha = a, ci_lo = ci[1], ci_hi = ci[2])
})
knitr::kable(bind_rows(alpha_rows), digits = 3)
```

## Approach labels: Jaccard (both conventions)

```{r jaccard}
pr <- unique(df[, c("Id", "anotator_name", "Reseni_gen_AI",
                    "Pristup.I", "Pristup.II", "Pristup.III")])
for (c_ in c("Pristup.I", "Pristup.II", "Pristup.III"))
  pr[[c_]][pr[[c_]] %in% c("", "nic")] <- NA
pr$item <- paste0(pr$Id, "_", pr$Reseni_gen_AI)
sets <- pr |> rowwise() |>
  mutate(s = list(na.omit(c(Pristup.I, Pristup.II, Pristup.III)))) |> ungroup() |>
  select(anotator_name, item, s) |> distinct(anotator_name, item, .keep_all = TRUE)
rat <- unique(sets$anotator_name)
jac <- function(sets) {
  inc <- c(); exc <- c()
  for (i in seq_along(rat)) for (j in seq_along(rat)) if (i < j) {
    a <- sets |> filter(anotator_name == rat[i])
    b <- sets |> filter(anotator_name == rat[j])
    m2 <- inner_join(a, b, by = "item", suffix = c("_a", "_b"))
    inc <- c(inc, mean(mapply(function(x, y) {
      if (!length(x) && !length(y)) return(1)
      length(intersect(x, y)) / length(union(x, y)) }, m2$s_a, m2$s_b)))
    exc <- c(exc, mean(mapply(function(x, y) {
      if (!length(x) && !length(y)) return(NA_real_)
      length(intersect(x, y)) / length(union(x, y)) }, m2$s_a, m2$s_b), na.rm = TRUE))
  }
  c(incl = mean(inc), excl = mean(exc))
}
j <- jac(sets)
put("jaccard_pristupy_obe_prazdne_shoda", j["incl"])
put("jaccard_pristupy_bez_prazdnych", j["excl"])
knitr::kable(data.frame(konvence = c("both-empty = agreement (paper)", "exclude both-empty"),
                        jaccard = round(j, 3)))
```

## Approach labels: prevalence and GLMM (RQ2 of the manuscript)

Binary outcome: any framework label assigned to a solution by a rater.
GLMM with logit link, random intercepts for case and annotator (as in the
manuscript's analyses.R; expected OR ≈ 0.31 for teacher solutions).

```{r glmm-frameworks}
pr2 <- pr |> rowwise() |>
  mutate(n_filled = sum(!is.na(c_across(starts_with("Pristup"))))) |> ungroup() |>
  group_by(Id, Reseni_gen_AI, anotator_name) |>
  summarise(n_filled = max(n_filled), .groups = "drop") |>
  mutate(filled_binary = as.integer(n_filled > 0))
n_labels <- sum(!is.na(pr$Pristup.I)) + sum(!is.na(pr$Pristup.II)) + sum(!is.na(pr$Pristup.III))
lab_by_src <- pr |> pivot_longer(starts_with("Pristup"), values_to = "lab") |>
  filter(!is.na(lab)) |> count(Reseni_gen_AI)
podil_AI <- lab_by_src$n[lab_by_src$Reseni_gen_AI == 0] / sum(lab_by_src$n)
gm <- lme4::glmer(filled_binary ~ Reseni_gen_AI + (1 | Id) + (1 | anotator_name),
                  data = pr2, family = binomial)
gs <- summary(gm)$coefficients
gci <- exp(confint(gm, parm = "beta_", method = "Wald"))
put("ramce_n_labelu_celkem", n_labels)
put("ramce_podil_labelu_AI", podil_AI)
put("glmm_or_ucitel_ramce", exp(gs[2, 1]))
put("glmm_or_ci_lo_ramce", gci[2, 1])
put("glmm_or_ci_hi_ramce", gci[2, 2])
knitr::kable(data.frame(
  n_labels_total = n_labels, share_AI = round(podil_AI, 3),
  OR_teacher = round(exp(gs[2, 1]), 3),
  ci = sprintf("[%.3f; %.3f]", gci[2, 1], gci[2, 2]), p = signif(gs[2, 4], 3)))
```

## Freeze: write manifest

```{r manifest}
out <- bind_rows(manifest) |>
  mutate(skript = "analyzy/notebooks/30_ai_vs_ucitel.qmd",
         data_sha = DATA_SHA, seed = SEED, datum = format(Sys.Date()))
dir.create(file.path(H2, "vystupy/tabulky"), showWarnings = FALSE, recursive = TRUE)
write.csv(out, file.path(H2, "vystupy/tabulky/kap7_k2_shoda_cisla.csv"), row.names = FALSE)
cat("Manifest:", nrow(out), "metrik → vystupy/tabulky/kap7_k2_shoda_cisla.csv\n")
sessionInfo()$R.version$version.string
```
