---
title: "E4 — Quality of case studies, reliability and K1 × K3 (Study 4, K3 campaign)"
author: "EduStories / habilitace-2"
date: today
format:
  html:
    embed-resources: true
    toc: true
execute:
  warning: false
  message: false
---

Frozen analysis of the third annotation campaign (K3) on the slice
`K3_freeze_2026-09-08` (PLAN.md §5 E4). Inputs are derived tables **without
case texts** (`data/processed/`): `k3_wide.csv` (one row = case × annotator,
A12 / A2), `k1_k3_pary.csv` (cases coded in K1 (2024) and K3 (2026)),
`anotace_pokryti.csv`, the July monitoring point
(`k3_monitoring_log.csv`), historical July and adjudication summaries
(`k3_cervenec_souhrn.csv`, `k3_adjudikace_souhrn.csv`) and reconstructed
segments (`k3_segmenty.csv`). These historical summaries are supplied inputs,
not a newly repeated annotation or source-text audit. Pre-specified protocol: Gwet AC1
next to κ/α; **primary reliability excludes adjudicated cases** (cases listed in
the July recalibration briefing), the full overlap is reported as sensitivity;
appropriateness K2 (1–5) and K3 (1–4) are never merged. Writes
`vystupy/tabulky/kap8_kvalita_cisla.csv` (+ `kap8_matice_zamen.csv`,
`kap8_reseni_dopad.csv`, `kap8_kvartily.csv`, `kap8_segmenty.csv`).

```{r setup}
suppressPackageStartupMessages({
  library(dplyr); library(tidyr); library(irr); library(irrCAC); library(digest)
})
SEED <- 20260714
set.seed(SEED)
B <- 1000
ROOT <- normalizePath(file.path(getwd(), ".."))          # analyzy/
H2   <- normalizePath(file.path(ROOT, ".."))             # habilitace-2/
PROC <- file.path(H2, "data/processed")
W  <- read.csv(file.path(PROC, "k3_wide.csv"), encoding = "UTF-8", stringsAsFactors = FALSE)
PK <- read.csv(file.path(PROC, "k1_k3_pary.csv"), encoding = "UTF-8", stringsAsFactors = FALSE)
input_names <- c("k3_wide.csv", "k1_k3_pary.csv", "k3_segmenty.csv",
                 "k3_monitoring_log.csv", "k3_cervenec_souhrn.csv", "k3_adjudikace_souhrn.csv")
input_hashes <- vapply(file.path(PROC, input_names), function(path) digest(file = path, algo = "sha256"), character(1))
DATA_SHA <- digest(paste(input_names, input_hashes, collapse = "\n"), algo = "sha256", serialize = FALSE)
W$case_uid[is.na(W$case_uid)] <- ""
W$kvalita[W$kvalita == ""] <- NA; W$dopad[W$dopad == ""] <- NA
stopifnot(all(W$anotator %in% c("A12", "A2")))

manifest <- list()
put <- function(metric, value) manifest[[length(manifest) + 1]] <<-
  data.frame(metric = metric, value = round(as.numeric(value), 4))
pct <- function(x) round(100 * x, 1)

# ── agreement helpers (two raters, one row per unit) ──────────────────────
exact <- function(a, b) mean(a == b)
kappa_nom <- function(a, b) irr::kappa2(cbind(a, b))$value
kappa_w  <- function(a, b) irr::kappa2(cbind(a, b), weight = "squared")$value
alpha_nom <- function(a, b) irr::kripp.alpha(rbind(as.integer(factor(c(a, b)))[seq_along(a)],
                                                   as.integer(factor(c(a, b)))[-seq_along(a)]),
                                             method = "nominal")$value
alpha_ord <- function(a, b) irr::kripp.alpha(rbind(a, b), method = "ordinal")$value
ac1 <- function(a, b) irrCAC::gwet.ac1.raw(data.frame(a, b))$est$coeff.val
ac2 <- function(a, b) irrCAC::gwet.ac1.raw(data.frame(a, b), weights = "quadratic")$est$coeff.val
boot_ci <- function(a, b, f, B = 1000) {
  # Seed per call: an interval must not depend on how many bootstraps ran earlier in
  # the notebook, otherwise adding a section silently changes every later interval.
  set.seed(SEED)
  n <- length(a); v <- replicate(B, { i <- sample.int(n, n, TRUE); f(a[i], b[i]) })
  quantile(v, c(.025, .975), na.rm = TRUE, names = FALSE)
}
```

## 1 Scope and coverage

```{r scope}
for (a in c("A12", "A2")) {
  wa <- W[W$anotator == a, ]
  put(paste0("n_radku_", a), nrow(wa))
  put(paste0("pokryti_kvalita_", a), mean(!is.na(wa$kvalita)))
  put(paste0("pokryti_dopad_", a), mean(!is.na(wa$dopad)))
  put(paste0("pokryti_vhodnost_", a), mean(!is.na(wa$vhodnost)))
  put(paste0("adjudikovano_radku_", a), sum(wa$adjudikovano))
}
uids <- unique(W$case_uid[W$case_uid != ""])
put("n_kazuistik_k3_unikat", length(uids))
put("n_kazuistik_k3_horni_mez", length(uids) + sum(W$case_uid == ""))
put("n_radku_bez_uid", sum(W$case_uid == ""))
ov <- W %>% filter(case_uid != "") %>% count(case_uid) %>% filter(n == 2)
put("n_prekryv", nrow(ov))
adj_uid <- unique(W$case_uid[W$adjudikovano == 1 & W$case_uid != ""])
put("n_prekryv_adjudikovano", sum(ov$case_uid %in% adj_uid))
knitr::kable(bind_rows(manifest))
```

## 2 Distributions per annotator

```{r distributions}
for (a in c("A12", "A2")) {
  wa <- W[W$anotator == a, ]
  tk <- table(wa$kvalita); td <- table(wa$dopad); tv <- table(wa$vhodnost)
  put(paste0("kvalita_plocha_n_", a), sum(wa$kvalita == "Plochá", na.rm = TRUE))
  put(paste0("kvalita_plocha_podil_", a), sum(wa$kvalita == "Plochá", na.rm = TRUE) / sum(!is.na(wa$kvalita)))
  put(paste0("kvalita_plna_n_", a), sum(wa$kvalita == "Plná", na.rm = TRUE))
  for (k in c("Neúspěch", "Krátkodobý úspěch", "Dlouhodobý úspěch")) {
    key <- c("Neúspěch" = "NEU", "Krátkodobý úspěch" = "KU", "Dlouhodobý úspěch" = "DU")[[k]]
    put(paste0("dopad_", key, "_n_", a), sum(wa$dopad == k, na.rm = TRUE))
    put(paste0("dopad_", key, "_podil_", a), sum(wa$dopad == k, na.rm = TRUE) / sum(!is.na(wa$dopad)))
  }
  for (v in 1:4) {
    put(paste0("vhodnost_", v, "_n_", a), sum(wa$vhodnost == v, na.rm = TRUE))
    put(paste0("vhodnost_", v, "_podil_", a), sum(wa$vhodnost == v, na.rm = TRUE) / sum(!is.na(wa$vhodnost)))
  }
  put(paste0("vhodnost_prumer_", a), mean(wa$vhodnost, na.rm = TRUE))
  pl <- wa[wa$kvalita %in% "Plochá", ]; pn <- wa[wa$kvalita %in% "Plná", ]
  put(paste0("plocha_bez_dopadu_", a), sum(is.na(pl$dopad)))
  put(paste0("plocha_bez_vhodnosti_", a), sum(is.na(pl$vhodnost)))
  put(paste0("plna_bez_dopadu_", a), sum(is.na(pn$dopad)))
  put(paste0("chybi_dopad_", a), sum(is.na(wa$dopad)))
  # comments and text lengths
  put(paste0("komentar_jazykovy_n_", a), sum(wa$komentar_jazykovy))
  put(paste0("komentar_plocha_n_", a), sum(wa$komentar_plocha))
  put(paste0("komentar_dalsi_n_", a), sum(wa$komentar_dalsi))
  put(paste0("otazniky_n_", a), sum(wa$otazniky))
  put(paste0("tema_strojovy_preklad_n_", a), sum(wa$tema_strojovy_preklad))
  put(paste0("plocha_s_komentarem_", a), sum(pl$komentar_plocha))
}
w12 <- W[W$anotator == "A12", ]
for (p in c("popis", "anamneza", "reseni", "vystup"))
  put(paste0("median_znaku_", p), median(w12[[paste0("n_znaku_", p)]]))
```

## 3 Reliability on the double-coded overlap

Primary set = overlap **without** cases discussed at the July recalibration
meeting (they are not independent double measurements); sensitivity = full
overlap. Bootstrap over cases, B = `r B`, seed `r SEED`.

```{r reliability}
P <- W %>% filter(case_uid %in% ov$case_uid) %>%
  select(case_uid, anotator, kvalita, dopad, vhodnost, adjudikovano) %>%
  pivot_wider(id_cols = case_uid, names_from = anotator,
              values_from = c(kvalita, dopad, vhodnost, adjudikovano))
P$adj <- pmax(P$adjudikovano_A12, P$adjudikovano_A2)
P$poradi <- W$poradi[match(P$case_uid, W$case_uid[W$anotator == "A12"])]
prim <- P[P$adj == 0, ]
rel <- function(D, prefix, dimensions = c("kvalita", "dopad", "vhodnost")) {
  # kvalita, dopad: nominal; vhodnost: ordinal
  for (dim in dimensions) {
    a <- D[[paste0(dim, "_A12")]]; b <- D[[paste0(dim, "_A2")]]
    ok <- !is.na(a) & !is.na(b); a <- a[ok]; b <- b[ok]
    key <- paste0(prefix, dim)
    put(paste0(key, "_n"), length(a))
    put(paste0(key, "_presna"), exact(a, b))
    ci <- boot_ci(a, b, exact, B); put(paste0(key, "_presna_ci_lo"), ci[1]); put(paste0(key, "_presna_ci_hi"), ci[2])
    if (dim == "vhodnost") {
      put(paste0(key, "_pm1"), mean(abs(a - b) <= 1))
      put(paste0(key, "_wkappa"), kappa_w(a, b))
      ci <- boot_ci(a, b, kappa_w, B); put(paste0(key, "_wkappa_ci_lo"), ci[1]); put(paste0(key, "_wkappa_ci_hi"), ci[2])
      put(paste0(key, "_alpha_ord"), alpha_ord(a, b))
      ci <- boot_ci(a, b, alpha_ord, B); put(paste0(key, "_alpha_ord_ci_lo"), ci[1]); put(paste0(key, "_alpha_ord_ci_hi"), ci[2])
      put(paste0(key, "_ac1"), ac1(a, b)); put(paste0(key, "_ac2"), ac2(a, b))
      ci <- boot_ci(a, b, ac1, B); put(paste0(key, "_ac1_ci_lo"), ci[1]); put(paste0(key, "_ac1_ci_hi"), ci[2])
      put(paste0(key, "_A2_vyssi"), sum(b > a)); put(paste0(key, "_A12_vyssi"), sum(a > b))
      put(paste0(key, "_prumer_A12"), mean(a)); put(paste0(key, "_prumer_A2"), mean(b))
      put(paste0(key, "_rozdil_prumer"), mean(b - a))
      ci <- boot_ci(a, b, function(x, y) mean(y - x), B)
      put(paste0(key, "_rozdil_ci_lo"), ci[1]); put(paste0(key, "_rozdil_ci_hi"), ci[2])
    } else {
      put(paste0(key, "_kappa"), kappa_nom(a, b))
      ci <- boot_ci(a, b, kappa_nom, B); put(paste0(key, "_kappa_ci_lo"), ci[1]); put(paste0(key, "_kappa_ci_hi"), ci[2])
      put(paste0(key, "_alpha"), alpha_nom(a, b))
      ci <- boot_ci(a, b, alpha_nom, B); put(paste0(key, "_alpha_ci_lo"), ci[1]); put(paste0(key, "_alpha_ci_hi"), ci[2])
      put(paste0(key, "_ac1"), ac1(a, b))
      ci <- boot_ci(a, b, ac1, B); put(paste0(key, "_ac1_ci_lo"), ci[1]); put(paste0(key, "_ac1_ci_hi"), ci[2])
    }
  }
}
rel(prim, "shoda_")   # primary (without adjudicated)
rel(P, "plny_")       # sensitivity (full overlap)
complete_quality <- prim[prim$kvalita_A12 == "Plná" & prim$kvalita_A2 == "Plná", ]
rel(complete_quality, "obe_plne_", dimensions = c("dopad", "vhodnost"))
for (annotator in c("A12", "A2")) {
  flat_records <- W[W$anotator == annotator & W$kvalita %in% "Plochá", ]
  put(paste0("plocha_s_dopadem_", annotator), sum(!is.na(flat_records$dopad)))
  put(paste0("plocha_s_vhodnosti_", annotator), sum(!is.na(flat_records$vhodnost)))
}

# confusion matrices (full overlap; A12 rows, A2 columns)
mz <- bind_rows(lapply(c("kvalita", "dopad", "vhodnost"), function(dim) {
  a <- P[[paste0(dim, "_A12")]]; b <- P[[paste0(dim, "_A2")]]
  ok <- !is.na(a) & !is.na(b)
  as.data.frame(table(A12 = a[ok], A2 = b[ok])) %>% mutate(dimenze = dim) %>%
    rename(n = Freq) %>% select(dimenze, A12, A2, n)
}))
write.csv(mz, file.path(H2, "vystupy/tabulky/kap8_matice_zamen.csv"), row.names = FALSE)
rec <- bind_rows(manifest); knitr::kable(rec[grepl("^shoda_|^plny_", rec$metric), ])
```

## 4 Recalibration: July monitoring point and order quartiles

```{r drift}
log <- file.path(PROC, "k3_monitoring_log.csv")
stopifnot(file.exists(log))
if (file.exists(log)) {
  m <- read.csv(log, encoding = "UTF-8"); m <- m[m$datum == "2026-07-16", ]
  put("cervenec_prekryv_n", m$prekryv_hotovo)
  put("cervenec_vhodnost_n", m$vhodnost_n); put("cervenec_vhodnost_presna", m$vhodnost_shoda)
  put("cervenec_vhodnost_wkappa", m$vhodnost_koef); put("cervenec_vhodnost_ac1", m$vhodnost_ac1)
  put("cervenec_dopad_n", m$dopad_n); put("cervenec_dopad_presna", m$dopad_shoda); put("cervenec_dopad_kappa", m$dopad_koef)
  put("cervenec_kvalita_n", m$kvalita_n); put("cervenec_kvalita_presna", m$kvalita_shoda); put("cervenec_kvalita_ac1", m$kvalita_ac1)
}
# direction of the July drift from the recalibration briefing (text parsed, no case texts)
july_summary <- read.csv(file.path(PROC, "k3_cervenec_souhrn.csv"))
for (summary_row in seq_len(nrow(july_summary)))
  put(july_summary$metric[summary_row], july_summary$value[summary_row])
# quartiles of the full overlap by A12 processing order (exploratory)
Q <- P %>% arrange(poradi) %>% mutate(kvartil = ntile(poradi, 4))
kv <- bind_rows(lapply(1:4, function(q) {
  d <- Q[Q$kvartil == q, ]; out <- data.frame(kvartil = q, n = nrow(d))
  for (dim in c("kvalita", "dopad", "vhodnost")) {
    a <- d[[paste0(dim, "_A12")]]; b <- d[[paste0(dim, "_A2")]]; ok <- !is.na(a) & !is.na(b)
    out[[paste0(dim, "_presna")]] <- exact(a[ok], b[ok])
    put(paste0("kvartil", q, "_", dim, "_presna"), exact(a[ok], b[ok]))
  }
  out
}))
write.csv(kv, file.path(H2, "vystupy/tabulky/kap8_kvartily.csv"), row.names = FALSE)
# Segments before/after the recalibration, reconstructed by 28_k3_segmenty.py from the
# July snapshot of both tables (backup of 2026-07-22 = state before the meeting; the
# script verifies it reproduces the July monitoring point exactly).
#   beze_zmeny = rated by 22 July and unchanged since -> BEFORE the recalibration
#   jen_finale = rated only after 22 July              -> AFTER the recalibration
#   zmeneno    = rated in July and rewritten afterwards (retroactive revision)
seg_f <- file.path(PROC, "k3_segmenty.csv")
if (file.exists(seg_f)) {
  SG <- read.csv(seg_f, encoding = "UTF-8")
  put("segmenty_dostupne", 1)
  put("seg_beze_zmeny_n", sum(SG$segment == "beze_zmeny"))
  put("seg_jen_finale_n", sum(SG$segment == "jen_finale"))
  put("seg_zmeneno_n", sum(SG$segment == "zmeneno"))
  put("seg_zmeneno_A12", sum(SG$segment == "zmeneno" & SG$anotator == "A12"))
  put("seg_zmeneno_A2", sum(SG$segment == "zmeneno" & SG$anotator == "A2"))
  for (dim in c("kvalita", "dopad", "vhodnost"))
    put(paste0("seg_zmeneno_", dim), sum(SG[[paste0("zmena_", dim)]], na.rm = TRUE))
  put("seg_zmeneno_projednano", sum(SG$segment == "zmeneno" & SG$adjudikovano == 1))
  put("seg_cervenec_radku_A12", sum(SG$anotator == "A12" & SG$segment != "jen_finale"))
  put("seg_cervenec_radku_A2", sum(SG$anotator == "A2" & SG$segment != "jen_finale"))
  s12 <- SG[SG$anotator == "A12", c("case_uid", "segment")]; names(s12)[2] <- "seg_A12"
  s2  <- SG[SG$anotator == "A2",  c("case_uid", "segment")]; names(s2)[2]  <- "seg_A2"
  PS <- P %>% left_join(s12, by = "case_uid") %>% left_join(s2, by = "case_uid")
  put("seg_prekryv_A2_cervenec", sum(PS$seg_A2 %in% c("beze_zmeny", "zmeneno"), na.rm = TRUE))
  put("seg_prekryv_A2_po", sum(PS$seg_A2 == "jen_finale", na.rm = TRUE))
  put("seg_prekryv_A12_po", sum(PS$seg_A12 == "jen_finale", na.rm = TRUE))
  put("seg_prekryv_A12_zmeneno", sum(PS$seg_A12 == "zmeneno", na.rm = TRUE))
  ok <- !is.na(PS$seg_A12) & !is.na(PS$seg_A2)
  PRED <- PS[ok & PS$seg_A12 == "beze_zmeny" & PS$seg_A2 == "beze_zmeny", ]
  PO   <- PS[ok & PS$seg_A12 == "jen_finale" & PS$seg_A2 == "beze_zmeny", ]
  put("seg_pred_n", nrow(PRED)); put("seg_po_n", nrow(PO))
  if (nrow(PRED) > 20) rel(PRED, "segpred_")
  if (nrow(PO) > 20) rel(PO, "segpo_")
  write.csv(SG %>% count(anotator, segment),
            file.path(H2, "vystupy/tabulky/kap8_segmenty.csv"), row.names = FALSE)
}
# adjudication summary (counts only; produced by 27_k3_freeze.py from the briefing sheet)
adjs <- file.path(PROC, "k3_adjudikace_souhrn.csv")
stopifnot(file.exists(adjs))
adjudication_summary <- read.csv(adjs)
for (summary_row in seq_len(nrow(adjudication_summary)))
  put(adjudication_summary$metric[summary_row], adjudication_summary$value[summary_row])
knitr::kable(kv, digits = 3)
```

## 5 K1 × K3: replication of outcome and solution type vs outcome

```{r k1k3}
put("k1k3_n_par", nrow(PK))
put("k1k3_kvalita_flat_k1", sum(PK$k1_flat == "flat", na.rm = TRUE))
# outcome K1 (merged to 3 categories) vs K3, unit = case × K3 annotator
long <- bind_rows(lapply(c("A12", "A2"), function(a)
  data.frame(case_uid = PK$case_uid, k1 = PK$k1_dopad3, k3 = PK[[paste0("k3_dopad_", a)]],
             vh = PK[[paste0("k3_vhodnost_", a)]], res = PK$k1_reseni_primarni, stringsAsFactors = FALSE)))
d <- long[!is.na(long$k1) & long$k1 != "" & !is.na(long$k3) & long$k3 != "", ]
put("k1k3_neuplny_k1_n", sum((is.na(long$k1) | long$k1 == "") & !is.na(long$k3) & long$k3 != ""))
put("k1k3_dopad_n", nrow(d)); put("k1k3_dopad_presna", exact(d$k1, d$k3))
put("k1k3_dopad_kappa", kappa_nom(d$k1, d$k3)); put("k1k3_dopad_ac1", ac1(d$k1, d$k3))
# cluster bootstrap over cases
cb <- function(f) { set.seed(SEED); u <- unique(d$case_uid); v <- replicate(B, {
  s <- sample(u, length(u), TRUE); dd <- d[unlist(lapply(s, function(x) which(d$case_uid == x))), ]
  f(dd$k1, dd$k3) }); quantile(v, c(.025, .975), names = FALSE) }
ci <- cb(exact); put("k1k3_dopad_presna_ci_lo", ci[1]); put("k1k3_dopad_presna_ci_hi", ci[2])
ci <- cb(kappa_nom); put("k1k3_dopad_kappa_ci_lo", ci[1]); put("k1k3_dopad_kappa_ci_hi", ci[2])
mz2 <- as.data.frame(table(K1 = d$k1, K3 = d$k3)) %>% rename(n = Freq)
write.csv(mz2, file.path(H2, "vystupy/tabulky/kap8_matice_k1k3.csv"), row.names = FALSE)

# solution type (K1 primary) × K3 outcome / appropriateness
r <- long[!is.na(long$res) & long$res != "", ]
r$pair_complete <- !is.na(r$k1) & r$k1 != "" & !is.na(r$k3) & r$k3 != ""
r$du <- as.integer(r$k3 == "Dlouhodobý úspěch")
r$du[!r$pair_complete] <- NA
tab <- r %>% group_by(res) %>% summarise(n_kazuistik = n_distinct(case_uid), n = sum(!is.na(du)),
  du_podil = mean(du, na.rm = TRUE), vhodnost_prumer = mean(vh, na.rm = TRUE), .groups = "drop") %>%
  arrange(desc(n_kazuistik))
put("k1k3_reseni_n", n_distinct(r$case_uid))
cbres <- function(cat, f) { set.seed(SEED); u <- unique(r$case_uid[r$res == cat]); v <- replicate(B, {
  s <- sample(u, length(u), TRUE); dd <- r[r$res == cat, ][unlist(lapply(s, function(x) which(r$case_uid[r$res == cat] == x))), ]
  f(dd) }); quantile(v, c(.025, .975), names = FALSE, na.rm = TRUE) }
tab$du_ci_lo <- NA; tab$du_ci_hi <- NA; tab$vh_ci_lo <- NA; tab$vh_ci_hi <- NA
tab$du_podil_k1 <- NA; tab$n_k1 <- NA
slug <- function(x) { x <- tolower(iconv(x, to = "ASCII//TRANSLIT")); gsub("^_+|_+$", "", gsub("[^a-z0-9]+", "_", x)) }
for (i in seq_len(nrow(tab))) {
  cat_ <- tab$res[i]; s <- slug(cat_)
  ci <- cbres(cat_, function(dd) mean(dd$du, na.rm = TRUE)); tab$du_ci_lo[i] <- ci[1]; tab$du_ci_hi[i] <- ci[2]
  ci <- cbres(cat_, function(dd) mean(dd$vh, na.rm = TRUE)); tab$vh_ci_lo[i] <- ci[1]; tab$vh_ci_hi[i] <- ci[2]
  put(paste0("reseni_", s, "_n"), tab$n_kazuistik[i])
  put(paste0("reseni_", s, "_du_podil"), tab$du_podil[i])
  put(paste0("reseni_", s, "_du_ci_lo"), tab$du_ci_lo[i]); put(paste0("reseni_", s, "_du_ci_hi"), tab$du_ci_hi[i])
  put(paste0("reseni_", s, "_vhodnost_prumer"), tab$vhodnost_prumer[i])
  put(paste0("reseni_", s, "_vh_ci_lo"), tab$vh_ci_lo[i]); put(paste0("reseni_", s, "_vh_ci_hi"), tab$vh_ci_hi[i])
  # same cases, outcome as coded by the FIRST campaign: makes the two teams comparable
  k1sub <- r[r$res == cat_ & r$pair_complete, ]
  tab$du_podil_k1[i] <- mean(k1sub$k1 == "Dlouhodobý úspěch")
  tab$n_k1[i] <- nrow(k1sub)
  put(paste0("reseni_", s, "_du_podil_k1"), tab$du_podil_k1[i])
  put(paste0("reseni_", s, "_n_k1"), tab$n_k1[i])
}
ct <- table(r$res[r$pair_complete], r$k3[r$pair_complete])
chi <- suppressWarnings(chisq.test(ct))
put("k1k3_reseni_dopad_cramer_v", sqrt(chi$statistic / (sum(ct) * (min(dim(ct)) - 1))))
write.csv(tab, file.path(H2, "vystupy/tabulky/kap8_reseni_dopad.csv"), row.names = FALSE)
knitr::kable(tab, digits = 3)
```

## 6 Regression checks against the descriptive scripts (22/23) and freeze

```{r checks}
v <- function(k) rec$value[rec$metric == k]
rec <- bind_rows(manifest)
stopifnot(abs(v("plny_dopad_kappa") - 0.674) < 0.01,
          abs(v("plny_dopad_presna") - 0.801) < 0.01,
          abs(v("plny_vhodnost_wkappa") - 0.531) < 0.01,
          abs(v("plny_kvalita_presna") - 0.967) < 0.01,
          v("n_prekryv") >= 690)
stopifnot(!any(duplicated(rec$metric)))
out <- rec %>% mutate(skript = "analyzy/notebooks/40_kvalita.qmd", data_sha = DATA_SHA,
                      seed = SEED, datum = format(Sys.Date()))
write.csv(out, file.path(H2, "vystupy/tabulky/kap8_kvalita_cisla.csv"), row.names = FALSE)
cat("manifest:", nrow(out), "metrics →", "vystupy/tabulky/kap8_kvalita_cisla.csv\n")
```

## 7 Stored predictions: offline recalculation

These tables are recalculated by `29_predikce_offline.py` without API calls.
Invalid responses remain in the denominator as failed predictions. Intervals
are retrospective case bootstraps; paired differences do not establish model
equivalence. Human reading of the reported outcome is a different task.

```{r stored-predictions}
predictions <- read.csv(file.path(H2, "vystupy/tabulky/kap8_predikce_vysledky.csv"))
knitr::kable(predictions[, c("model", "varianta", "reference", "n", "n_invalid",
                            "coverage", "bal_acc", "bal_acc_ci_lo", "bal_acc_ci_hi",
                            "macro_f1", "kappa", "kappa_valid")], digits = 3)
differences <- read.csv(file.path(H2, "vystupy/tabulky/kap8_predikce_rozdily.csv"))
knitr::kable(differences, digits = 4)
knitr::kable(rec[grepl("^obe_plne_.*(n$|kappa$|presna$)", rec$metric), ])
sessionInfo()
```
